---
title: "Assignment 2"
author: "Iris van Engen, Sara Claus, Caspar Bannink, Wessel ??"
output: html_document
date: "2023-10-25"
---

Libraries:
```{r}
library(ggplot2)
library(tidyverse)
```


# Introducing the dataset
To do:  
- choose a dataset  
- clean dataset
- explain dataset
- introduce what question we want to answer (so what is the dependent variable and are the independent variables)

# Representing the data
Content: make few interesting plots

# Simple model
To do: 
- Make a simple model
- Explain model method (linear regression?)
- Create confusion matrix 
- Explain de predictions (which are the truly true, falsly true etc)
- Evaluate the performance of the model
- Interpret parameters and/or contribution of the features (what is the difference here?)

```{r}
# Function measuring the performance of the model (last two values are the most interesting!)

model_performance = function(model_cm){
  
  TP <- model_cm[4]
  TN <- model_cm[1]
  FP <- model_cm[2]
  FN <- model_cm[3]

  ACC <- (TP+TN)/(TP+TN+FP+FN)
  TPR <- TP/(TP+FN) # true positive rate OR sensitivity
  TNR <- TN/(TN+FP) # true negative rate OR specificity
  FPR <- FP/(FP+TN) # false positive rate
  PPV <- TP/(TP+FP) # positive predicted value
  NPV <- TN/(TN+FP) # negative predicted value
  
  performance <- c("ACC:", ACC, "TPR:", TPR, "TNR:", TNR, "FPR:", FPR, "PPV:", PPV, "NPV:", NPV)
  
  print(performance)
}
```

# Complex model
To do: 
- Make a complex model
- Explain model method (logistic regression, k-nearest neighbour, linear discriminant analysis, random Forest ?)
- Create confusion matrix 
- Explain de predictions (which are the truly true, falsly true etc)
- Evaluate the performance of the model
- Interpret parameters and/or contribution of the features (what is the difference here?)

# Conclusion
To do:
- what do the models predict (answer the question we ask ourself at the beginning)
- conclude which model is the best and why


